---
title: "Osa data exploration"
output:
  html_notebook: 
    toc: yes
    code_folding: hide
    theme: spacelab
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(knitr); 
opts_chunk$set(echo = F)
library(here)
source(
  here("analysis/checkVars.R"))
library(vegan) #adonis()?
```

# Results

## Shade, *biomass*, *richness*

```{r plots}
mainModel <- median ~ dist
plotResultsTbl <- plotVarsTbl %>%
  checkVars(.,
            mainModel)

plotResultsTbl; plotResultsTbl$graph 
#ggarrange(.[[1]], .[[2]]) #buggy
```

-   spg et al. vars at tree or plot level?
-   otherwise richness, spg appear signif

## Trees

```{r trees}
treeVarsTbl$varData[[1]]
mainModel1 <- median1 ~ dist
treeResultsTbl <- 
  treeVarsTbl %>%

#nonormalmetrics #paused
treeResultsTbl1 <- treeResultsTbl %>%
  mutate(hist = varData %>% modify(~ hist(.x$median, plot = F)),
         histTbl = hist %>%
           modify(~ tibble(value = .x$mids, f = .x$counts, d = varData$dist)),
         histgraphs = varData %>%
           modify(~ .x %>% ggplot(aes(x = median)) +
                    stat_bin(geom = "point"),
         probs = hists %>% modify(~ disfit(.x$counts + 1))) #1min/variable
treeResultsTbl1$histgraphs
treeResultsTbl1$probs

treeResultsTbl; treeResultsTbl$graph #ggarrange(.[[1]], .[[2]]) #buggy
plot(treeResultsTbl$hists[[1]]$mids, treeResultsTbl$hists[[1]]$counts)
```

-   spg appears signif--explanation will be relevant
-   explore historgrams for non-sig vars? (i.e. non-equilibrium stats)

```{r biomass}

```

## Taxa

```{r entropy}
treeVarsTbl %>%
  mutate(spTbl = varData %>% 
           modify(~ .x %>% pivot_wider(names_from = gen, values_from = median)))
```

# Discussion

-   spg and richness change likely reflects shift in community (--\> add NMDS/PCoA, with closer plots dominated by pioneering species in dense stands
-   i think inference about dispersal strategies could be in discussion, or later figure, no need to be oriiginally embedded in analysis -- but would probably make story more ecological, and study more unique to presen, though
