---
title: "Osa data exploration"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    code_folding: hide
    theme: spacelab
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  collapse = T
  )
library(here)
source(
  here("analysis/checkVars.R")
  )
```

# Results

## Shade, *biomass*, *richness*

```{r get}
mainModel <- median ~ dist
plotResultsTbl <- plotVarsTbl %>%
  getStats(
    mainModel
  )
```


```{r table}
plotSignifTbl <- plotResultsTbl %>%
  select(
    variable,
    isSignif,
    isSignif1,
    graph
  ) %>%
  filter(
    isSignif == T |
      isSignif1 == T
  )
```

```{r plots}
plotSignifTbl %>%
  pull(
    graph
    )
#fixXaxis
#ggarrange(.[[1]], .[[2]]) #buggy
```


-   spg et al. vars at tree or plot level?
-   richness, entropy [+spg] signif
- but why reverse relationships?

## Trees

```{r trees}
mainModel1 <- median1 ~ dist
#treeVarsTbl %>%
```

-   spg appears signif--explanation will be relevant
-   explore historgrams for non-sig vars? (i.e. non-equilibrium stats)

```{r biomass}

```

## Taxa

# Discussion

-   spg and richness change likely reflects shift in community (--\> add NMDS/PCoA, with closer plots dominated by pioneering species in dense stands
-   i think inference about dispersal strategies could be in discussion, or later figure, no need to be oriiginally embedded in analysis -- but would probably make story more ecological, and study more unique to present, though
